Deep Learning for 3D Brain Activity Scans

### 2019-12-03 paper/sections/approach.tex
Final paper edits

Beginning with SKLearn~\cite{scikit-learn} data readiness was proven with moderately successful classification using
a simple SVM and then the SKLearn Multi-Layer-Perceptron neural network, getting 70-80\% accuracy, sufficing the
initial sanity check that deep learning would be ready to work on.

>>>>

Beginning with SKLearn~\cite{scikit-learn} data readiness was proven with moderately successful classification using
a simple SVM and then the SKLearn Multi-Layer-Perceptron neural network, getting 70-80\% accuracy, well above the
random chance accuracy of 50\%, sufficing the
initial sanity check that deep learning would be ready to work on.

### 2019-12-03 paper/sections/discussion.tex
Final paper edits

\subsection{Learning}\label{subsec:learning}
very cool! I tried spot checking different scans to see if I could tell the difference visually and I really really
cannot. I think it's encouraging that the model is able to notice impossible to visualize differences and run with it
this well. It's one thing to differentiate a st. bernard from a siamese cat, but two brain activations I'm not good at

\subsection{Future Works}\label{subsec:future-works}
Unsupervised segmentation, especially in 3d would be cool
Future works
- could strive to refine 3d more, with more time and compute
- unsupervised segmentation would be cool
- find ways to not have to smoosh images for memory?
- more visualizations? lack of color kinda sucks, saliency is weird, filters prob too

began work on visualization, it didn't lead anywhere useful, 1 channel might be a problem?
- check out filter viz
- see about unsupervised
- try 3d alternative archs? (dont really need to)

might want to improve 3d before going on to segmentation,

 \begin{figure}
  \includegraphics[width=\linewidth]{images/saliency.png}
  \caption{caption}
  \label{fig:saliency}
\end{figure}

>>>>

\subsection{Learning}\label{subsec:learning}

Using real world brain scans of a very low resolution, we've been able to achieve excellent classification using both
2 dimensional and 3 dimensional convolutional networks, with somewhat promising generalizability to previously unseen subjects.
The models have been very able to differentiate when a subject was scanned viewing a tempting food image against when the image
presented is neutral, where I personally can see absolutely no clear difference.

Unsurprisingly deeper architectures also tend to improve performance marginally,
as they can cope with increasingly subtle changes presumably, in both the performance of the 2d and 3d CNN's

While we may have empirically found improved performance using 3 dimensional CNN's over all alternatives,
the hefty cost of computation has proven difficult to cope with and limited the ability to explore segmentation farther.

\subsection{Future Works}\label{subsec:future-works}

While classification feels well proven and a desirable performance was achieved,
the natural next step is to try an highlight what the network has learned to recognize
as indicative of a particular stimuli, turning the blackbox computation into a powerful teacher.
My attempts made to the point of this submission to adapt the classifiers to unsupervised segmentation
have not yet been fully successful.

\subsubsection{Refinements}\label{refinements}

Refinements to the classifier are always a worthwhile endeavor, though this particular classification feels well
handled by the chosen architectures at relatively shallow levels.

Given the promise of warm-starts progressing learning, I'd be very interested in exploring different
learning rate schedulers to see what combination of minute and aggressive changes lands the model in the most performant
optimization.

Lastly as there's typically a free boost from using an ensemble model, I'd be interested in exploring what combination
of models yields a more robust cohesive model which generalizes very well to previously unseen subject's brain scans.

\subsubsection{Visualizations}\label{fw-viz}

The models are also naturally capable of producing visualizations of their convolutional filters, class representations
and creating saliency maps.

In figure~\ref{fig:saliency} we see one example of Saliency maps in action.
Interestingly enough the model does not key strongly off of discrete brain structures as I had assumed it would.
I speculate that the data being so downgraded provides difficulty in precisely conveying brain activity and the models
are instead forced to learn off of other subtle activities across the brain, which are of course not fully silent during
the passive image viewing task.

 \begin{figure}
  \includegraphics[width=\linewidth]{images/saliency.png}
  \caption{Three Nonfood \& Three Food Brain Scan Slices with their respective Saliency Maps}
  \label{fig:saliency}
\end{figure}

### 2019-12-03 paper/sections/results.tex
Final paper edits

the project explored minimally the hyperparameters of the chosen model architectures
chiefly was finding the right learning rate for the given normalized data to the model
typically 1e-2 to 1e-4 worked, with learning rates on the order of 1e-3 being best.
Stochastic Gradient Descent with nesterov momentum on Cross Entropy Loss performed very well.
Other optimizers like Adam and loss metrics like NLL Loss explored were either on par or worse and thus not
further utilized.

None of my attempted customizations to the ResNet architecture netted any improvements.

>>>>

I explored the available hyperparameters of the chosen model architectures to a minor degree, focusing on
establishing optimal learning rates with the chosen optimizer and model, with the given normalized data.

Typically 1e-2 to 1e-4 worked, with learning rates on the order of 1e-3 being best for both ResNet style convolutional
networks.

Stochastic Gradient Descent with nesterov momentum on Cross Entropy Loss performed very well.
Other optimizers like Adam and loss metrics like NLL Loss explored were either on par or worse and thus not
further utilized.

None of my attempted customizations to the ResNet architecture netted any improvements, including expanding filter sizes
or adding normalization layers.

### 2019-12-03 paper/submission.tex
Final paper edits

\author{Anonymized for Peer Review\\
University of Massachusetts Amherst\\
140 Governors Dr, Amherst, MA 01002\\
{\tt\small @umass.edu}
}

@@

% ------------------------------------------------------------------------
% ABSTRACT
%-------------------------------------------------------------------------
\begin{abstract}
 Analyzing the brain's activity patterns starts with identifying known patterns of well established structures,
 typically observed by human oversight in two dimensions, which may miss subtle, barely visible but useful indicators,
 and limit the ability to stitch together patterns across a third dimension.

@@

 \begin{figure}
  \includegraphics[width=\linewidth]{images/var_on_cortex_2.png}
  \caption{caption}
  \label{fig:cortex-var}
\end{figure}

>>>>

\author{Anonymous1\\
University of Massachusetts Amherst\\
140 Governors Dr, Amherst, MA 01002\\
{\tt\small @umass.edu}
}

@@

% ------------------------------------------------------------------------
% ABSTRACT
%-------------------------------------------------------------------------
\begin{abstract}
 Analyzing the brain's activity patterns starts with identifying known signals of well established structures,
 typically observed by human oversight in two dimensions, which may miss subtle, barely visible but useful indicators,
 and limit the ability to stitch together patterns across a third dimension.

@@

 \begin{figure}
  \includegraphics[width=\linewidth]{images/var_on_cortex_2.png}
  \caption{Region of Variance Across the Dataset of Brain Scans}
  \label{fig:cortex-var}
\end{figure}
